Spatial Analysis

### 2021-08-05 Scripts/spatial_analisis.Rmd
Created spatial polygon of ports data

---
title: "Spatial Analysis"
author: "Juliano Palacios Abrantes"
output:
  # html_document
  html_notebook:
    fig_width: 6
    fig_height: 4
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
    theme: darkly
---

@@

# Fix new updats of sf package
sf::sf_use_s2(use_s2 = FALSE)

@@

# Get numner of ports per state
ports_n_data <- port_sf %>% 
  as.data.frame() %>% 
    filter(distance_from_shore_m == 0) %>% 
    group_by(state) %>% 
    summarise(n_ports = n())

@@

# Get the overlapping segments 
# NOTE: Takes some gooooood time (~20 min)
Sys.time()
grid_eez_fp_sw <- st_intersection(grid_fp_sf,eez_sf)
Sys.time()
# write_sf(grid_eez_sf, paste0(my_path("D", "Spatial"),"grid_eez_sf.shp"))

# Get final df with index
grid_eez_fp_df <- as.data.frame(grid_eez_fp_sw) %>%
  select(state,index) %>%
  left_join(ne_grid,
            by = "index")

# Plot to make sure makes sense
grid_eez_fp_sw %>%
  ggplot() +
  geom_sf(aes(color = state), alpha = 0.3) +
  scale_color_manual(values = state_pallet) +
  geom_sf(data = eez_sf,aes(), fill = NA) +
  theme_dark()

@@

#### 3.3 Differences at cropede level

@@

```{r state_bf_diff_tbl, eval = T, echo = T}

@@

##### 3.1 Paper figure; buffer plot

@@

plot_data <-grid_eez_sf %>% 
    mutate(state = str_to_sentence(state)) %>% 
    left_join(state_lat)

p <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(plot_data) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color = order, fill = order), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                      size = 2,
                      box.padding = 0.10,
                      hjust = 1) +
    scale_color_manual(values = state_pallet) +
    scale_fill_manual(values = state_pallet) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 12) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45))+
    labs(x = "", y = ""),
  # Showing each state separately

    ggplot(plot_data) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    geom_sf( aes(color = order),size = 0.1, alpha = 0.8) +
    facet_wrap(~ state) +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet,
                       labels = plot_data %>% arrange(order) %>%  pull(state) %>% unique()) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 6,
                 axx_tx_ang = 45,
                 hjust = 1
                 ),
  nrow = 1,
  bottom = "Longitude",
  left = "Latitude")

  ggsave(filename = "buffer_figure.jpg",
         plot = p,
         path = my_path("R","Figures"),
         width = 10,
         height = 5
         )

@@

![Buffer figure for paper](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/buffer_figure.jpg)

@@

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_read, fread)) %>% 
  mutate(state = str_to_sentence(state))

>>>>

---
title: "Spatial Analysis"
author: "Juliano Palacios Abrantes"
output:
  html_document:
  # html_notebook:
    fig_width: 6
    fig_height: 4
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
    theme: darkly
---

@@

# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

@@

# Get number of ports per state
ports_n_data <- port_sf %>% 
  as.data.frame() %>% 
    filter(distance_from_shore_m == 0) %>% 
    group_by(state) %>% 
    summarise(n_ports = n())

@@

# Get the overlapping segments 
# NOTE: Takes some gooooood time (~20 min)
Sys.time()
grid_eez_sf_fp <- st_intersection(grid_fp_sf,eez_sf)
Sys.time()

# write_sf(grid_eez_sf_fp, paste0(my_path("D", "Spatial/grid_eez_fp_sf"),"grid_eez_fp.shp"))

# Get final df with index
grid_eez_fp_df <- as.data.frame(grid_eez_sf_fp) %>%
  select(state,index) %>%
  left_join(ne_grid,
            by = "index")

write_csv(grid_eez_fp_df, paste0(my_path("R", "Partial"),"grid_eez_fp_df.csv"))

# Plot to make sure makes sense
grid_eez_sf_fp %>%
  ggplot() +
  geom_sf(aes(color = state), alpha = 0.3) +
  scale_color_manual(values = state_pallet) +
  geom_sf(data = eez_sf,aes(), fill = NA) +
  theme_dark()

@@

#### 3.3 Differences at croped level

@@

```{r state_bf_diff_ntbl, eval = T, echo = T}

@@

#### 3.4 Paper plot; Buffer plot

@@

#### Data needed ####

# Get state order from North to south
state_order <- my_path("R","Partial/", name = "grid_eez_fp_df.csv",read = T) %>%
  group_by(state) %>% 
  summarise(
    order = mean(lat)
  ) %>% 
  mutate(abrev = c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")
  ) %>% 
  arrange(desc(order)) %>% 
  mutate(order= c("b","a",letters[3:11])) # weirdly maine goes below

# Load grid of state waters
grid_eez_sw_sf <- st_read(my_path("D","Spatial/grid_eez_sw_sf", name = "grid_eez_sf.shp")) %>% 
  select(index,state,geometry) %>% 
  mutate(method = "state_waters",
         state = str_to_sentence(state)) %>% 
  left_join(state_order)

grid_eez_sw_sf$group <- factor(grid_eez_sw_sf$abrev,      # Reordering group factor levels
                               levels = state_order$abrev)

# Load grid of fishing ports
grid_eez_fp_sf <- st_read(my_path("D","Spatial/grid_eez_fp_sf", name = "grid_eez_fp.shp")) %>% 
  select(index,state = jursdct,geometry) %>% 
  mutate(method = "fishing_port",
         state = str_to_sentence(state))%>% 
  left_join(state_order)

grid_eez_fp_sf$group <- factor(grid_eez_fp_sf$abrev,      # Reordering group factor levels
                               levels = state_order$abrev)

# Plotting 
# Make sure st_simplify() is run for eez sf

# State waters plot

p_sw <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_eez_sw_sf) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color =abrev , fill = abrev), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    scale_color_manual(values = state_pallet) +
    scale_fill_manual(values = state_pallet) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45))+
    labs(x = "", y = "", title = "A) State waters extrapolation") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot(grid_eez_sw_sf) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    geom_sf(aes(color = abrev),size = 0.1, alpha = 0.3) +
    facet_wrap(~ group) +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet,
                       # labels = unique(grid_eez_sw_sf$group)) +
                       labels = grid_eez_sw_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

##### ------------- ####
# Fishing Ports plot
##### ------------- ####

p_fp <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_eez_fp_sf) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color = abrev, fill = abrev), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    scale_color_manual(values = state_pallet) +
    scale_fill_manual(values = state_pallet) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45)) +
    labs(x = "", y = "", title = "B) Fishing ports extrapolation") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot(grid_eez_fp_sf) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    geom_sf(aes(color = abrev),size = 0.1, alpha = 0.3) +
    facet_wrap(~ group) +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet,
                       # labels = unique(grid_eez_fp_sf$group)) +
                       labels = grid_eez_fp_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

combined_plot <- gridExtra::grid.arrange(p_sw,p_fp,
                                         bottom = "Longitude",
                                         left = "Latitude")

ggsave(filename = "buffer_figure_b.jpg",
       plot = combined_plot,
       path = my_path("R","Figures"),
       width = 11,
       height = 11
)

@@

![Buffer figure for paper](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/buffer_figure_b.jpg)

@@

# Load all sp in one table
historic_spp_sw <- bind_rows(lapply(taxon_read, fread)) %>% 
  mutate(state = str_to_sentence(state))
